Factors influencing Total Income in the United States

### 2024-11-30 paper/paper.qmd
Modify paper

# Introduction
[] for citation 
appendix 
WEIGHT 
The weight (X42001) is a partially design-based
weight constructed at the Federal Reserve using original selection
probabilities and frame information along with aggregate control
totals estimated from the Current Population Survey.  This weight is a
relatively minor revision of the consistent weight series (X42000)
maintained for the SCFs beginning with 1989 (For a detailed discussion
of these weights, see "Consistent Weight Design for the 1989, 1992,
and 1995 SCFs and the Distribution of Wealth," by Arthur B. Kennickell
and R. Louise Woodburn, Review of Income and Wealth, Series 45, Number
2, June 1999, pp. 193-215 or the longer version given on the SCF web
site at www.federalreserve.gov/econresdata/scf/scf_workingpapers.htm).
The nature of the revisions to the consistent weights is described in
"Revisions to the SCF Weighting Methodology: Accounting for
Race/Ethnicity and Homeownership" (Arthur B. Kennickell, January 1999,
www.federalreserve.gov/econresdata/ scf/scf_workingpapers.htm).
A version of the revised weight has been computed for all the surveys
beginning with 1989, and this variable has been added to the public
versions of the SCF data sets.  Users should be aware that the
population defined by the weights for *each implicate* (see above) is
128.6 million households: the sum of each of the weights over all
sample cases and imputation replicates is equal to five times the
number of households in the sample universe.

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

@@

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (), from .

@@

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

@@

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

@@

# Model

@@

## Model set-up

@@

#first_model <-readRDS(file = here::here("models/first_model.rds"))
```

@@

#modelsummary::modelsummary(list("First model" = first_model),statistic = "mad",fmt = 2)
```

>>>>

# Introduction

Income inequality is a central concern in economic research, reflecting disparities in access to opportunities, resources, and outcomes. In the United States, total income is influenced by a variety of factors, including personal, professional, and demographic characteristics. This paper explores the relationship between total income and predictors such as marital status, sex, educational level, labor force participation, and occupation. By employing statistical modeling techniques, we aim to uncover the extent to which these variables shape income levels and contribute to broader income disparities.

The estimand of this study is the expected total income for individuals based on their social, personal and occupational characteristics. Specifically, the analysis seeks to quantify how factors such as marital status, sex, educational attainment, participation in the labor force, and occupational category predict variations in total income across the U.S. population. This includes identifying both independent and among these variables.

Preliminary results indicate that educational attainment is one of the strongest predictors of total income, with higher levels of education yielding significantly greater earnings. Marital status also plays a notable role, with married individuals generally earning more than their single counterparts. Gender disparities in income persist, with men earning more on average than women across most occupational categories. Labor force participation and occupation further highlight critical differences, as full-time workers and those in higher-skilled professions consistently report higher incomes.

Understanding the drivers of income disparities is essential for designing policies that promote economic equity and mobility. By identifying how marital status, gender, education, labor force participation, and occupation contribute to income levels, this research offers actionable insights for policymakers, employers, and educators. Addressing these disparities not only fosters greater fairness but also enhances the overall economic productivity and social cohesion of the United States.

To do this, total income was analyzed in Section 2 @sec-data. The prediction from the model has show accuracy in lower total income group. The model has shown difficulties in predicting higher total income. Section 3 @sec-model has further explanation on the Bayesian Generalized Linear Model. Section 4 @sec-results show the coefficient beta from the model. Section 5 @sec-discussion focus on the weakness of this study and further possible improvements. Lastly, Section 6 @sec-appendix contains methodology to perform a survey to collect data for Total Income.

@@

## Overview {#sec-data-overview}

The Survey of Consumer Finances (SCF), conducted in 2022 covering the period 1989 to 2022 by the Board of Governors of the Federal Reserve System of the United States [@scfdata], provides a comprehensive snapshot of household financial conditions across the nation. This dataset includes detailed information related to total income, capturing key predictors such as sex, education level, occupation category, salary income, and income from other sources. By offering rich, granular data, the SCF serves as an invaluable resource for analyzing the factors influencing income distribution and economic disparities within the United States.

## Measurement {#sec-data-measurement}

The survey contained a panel element over 2 periods. Respondents to the 1983 survey were re-interviewed in 1986 and 1989. Respondents to the 2007 survey were re-interviewed in 2009. In order to ensure the representativeness of the study, respondents are selected randomly in order to attempt families from all economic strata.

## Outcome variables{#sec-data-outcome}

Histograph of the predicted variables, Total income.

@@

```{r}
#| label: fig-qq
#| fig-cap: Comparison of Income Distribution using QQ Plot
#| echo: false
#| warning: false
#| message: false

@@

## Predictor variables{#sec-data-predictor}

Predictors variables selected are Age, Sex, Marital Status, Education Level, Labour Force, Work Status, and Occupation category.

@@

# Model{#sec-model}

@@

## Model set-up{#sec-model-setup}

@@

first_model <-readRDS(file = here::here("models/bayesian_model_2.rds"))
model plot
```

@@

modelsummary::modelsummary(list("First model" = first_model),statistic = "mad",fmt = 2)
```
